Generative Adversarial Networks

### 2017-12-04 cs6735/project/report.tex
edited up to the algorithm

\section{Introduction}
Deep learning has made astounding progress within the past decade. Discriminative models have already surpassed the abilities of humans to recognize patterns within certain domains \citep{2014arXiv1404.7828S}. These successes can attributed to vast, high dimension datasets in conjunction with large neural networks using linear activation functions, dropout regularization techniques and backpropagation to update the parameters \citep{2014arXiv1406.2661G}. However, deep learning possesses much more ambitious goals within the realm of unsupervised learning. As humans, we are able to understand the world around us with tremendous precision. It is easy to underestimate the complexity of the data we process to accomplish this feat. Although progress within the machine learning field is rapidly advancing, computers still have limited understanding of the data the process. Generative models, especially deep generative models such as GANs, offer promising results towards this formidable challenge \citep{genmodelingopenai}.

\subsection{Generative Modeling}
Generative models learn the joint probability distribution $p(x,y)$ of an input $x$ and label $y$ whereas discriminative models directly learn the conditional probability $p(y|x)$. An example of this process can be seen in figure \ref{fig:pdf}. Generative models may be used as classifiers using Bayes rules to calculate the conditional probability $p(x,y)$ which can then be used to make predictions \citep{NIPS2001_2020}. Although, generative models offer several other applications due to their knowledge of the underlying data distribution. Depending on the model, knowledge of the probability distribution can be created both explicitly and implicitly \citep{Goodfellow-et-al-2016}. Those which do not directly model a probability distribution offer mechanisms which require implicit knowledge of the underlying distribution, such as geminating a sample from that distribution \citep{Goodfellow-et-al-2016}.

@@

The process of training a generative model is very similar to a discriminative model. Using large amounts of data collected from a specific domain, we train the generative model to generate data from that domain. As these models typically have fewer parameters than that the number of data samples, they are forced to internalize some representation of the data \citep{genmodelingopenai}. This process within the context of GANs is elaborated in section \ref{sed:latentspace}. Unlike supervised training, there exists no single desired output. The problem becomes defining a cost function which forces the generative model to produce data more like that of the target domain \citep{genmodelingopenai}.

\subsection{Generative Adversarial Networks}
Generative adversarial networks (GANs), first introduced in 2014 by Goodfellow \textit{et al.}, offer a new framework for estimating generative models with use of an adversarial process \citep{2014arXiv1406.2661G}. These models offer a clever approach to solving the aforementioned problem of explicitly defining a cost function. Rather than training a single model, a discriminator is introduced. These two networks are pitted against each other in a minimax, zero-sum game. The generative model $G$ attempts to produce data that resembles that of the training set while the discriminative model $D$ attempts to discriminate between a real and generated (fake) sample. To train this network, $G$ attempts to maximize the error rate of $D$ whereas $D$ attempts to minimize it. The goal of the GAN algorithm is to reduce the distance between the data and generated probability distributions. Backpropagation is used to update the parameters and to train each model. In the ideal case, $D$ is unable to distinguish the generated samples from the real samples and produces an error rate of $\frac{1}{2}$ \citep{2014arXiv1406.2661G}.

The adversarial process is often analogized to a counterfeiter trying to fool a detective \citep{2014arXiv1406.2661G}. The detective will receive both real and counterfeited money and will try and distinguish between the two. This leads to a competitive scenario where both parties gradually improve their abilities until the counterfeits are indistinguishable from the true currency.

\subsection{Motivations}
There exists several compelling reasons for studying generative modeling, especially GANs \citep{2017arXiv170100160G}. Generative models are extremely useful when your goal is understand the underlying data generating distribution. The ability to learn the joint probability distribution of high dimensional data is relevant to both applied math and engineering \citep{2017arXiv170100160G}. Furthermore, GAN promise many real world applications. Several of these opportunities are expanded upon in section \ref{sec:applications}.

\subsection{Related Work}
Until the advent of GANs, most deep generative models provided a parametric specification of an probability distribution function \citep{2014arXiv1406.2661G}. These models could be trained using maximum likelihood estimation which often required gradient estimation. The most successful of these models is the deep Boltzmann machine \citep{2014arXiv1406.2661G}. The difficulties associated with gradient estimation motivated the development of \textit{generative machines}, models which implicitly represent the joint probability distribution by generating samples from that distribution rather than explicitly rendering a probability distribution.

@@

\section{Background}
The goal of the GAN training process is to determine the discriminator and generator parameters such that the classifier error is minimized and maximized respectively \citep{2017arXiv171007035C}. GANs may use any differentiable functions for their underlying implementation; however, optimal results are typically achieved when neural networks are used \citep{2014arXiv1406.2661G}. We first define a vector $z$ which is sampled from a prior noise distribution $z~p_z(z)$. This vector becomes the argument of the generator $G(z;\theta_g)$ where $G$ is a differentiable function and $\theta_g$ are its parameters. Another differentiable function $D(x;\theta_d)$ is defined for the discriminator. This function takes as input a vector $x$ from either the output of $G$ or a sample from the training set and is parametrized by $\theta_d$. The output of the discriminator is a scalar number which represents the probability that the sample $x$ was derived from the training data. In this situation, there exists a minimax game, a scenario derived from game theory, where $D$ attempts to maximize the probability of assigning a correct label and $G$ attempts to minimize $\log(1-D(G(z)))$. The value function $V(D,G)$ becomes:

@@

This is a standard cross entropy cost function that is trained on two minibatches, one from the dataset and one from the generator. For this to become a zero-sum game, the cost function of the generator is simply the negation of the discriminator:

@@

This version of the generator loss function is particularly useful for theoretical analysis as shown in the original GAN paper \citep{2014arXiv1406.2661G}. Goodfellow demonstrated that the learning within the minimax game is identical to minimizing the Jensen-Shannon divergence and that each player will eventually converge if both players are able to directly update the function space as oppose to the parameters within the network \citep{2017arXiv170100160G}. However, in this minimax version, the loss function of the generator suffers from a vanishing gradient when the discriminator confidently rejects the generated data. The solution proposed by Goodfellow is to train the generator to maximize $log(D(G(z)))$ rather than minimize $log(1-D(G(z)))$ \citep{2014arXiv1406.2661G}. This function is heuristically motivated as to provide strong gradients when either the generator or discriminator is not functioning optimally. In this version, the generator attempts to maximize the log probability of a misclassification whereas the minimax version attempts to minimize the log probability of the discriminator being correct:

@@

A maximum likelihood cost is another option for training GANs \citep{2017arXiv170100160G}. It can be shown that that this cost function minimizes the Kullback-Leibler divergence between $p_{data}$ and $p_{model}$ instead of the Jensen-Shannon divergence. This cost estimator is useful as it allows for comparison with other generative models; however, it behaves similarly to the minimax cost function as seen in figure \ref{fig:CostFunc}.

@@

\subsection{Algorithm}
This following algorithm presents initial training algorithm used by Goodfellow \textit{et al.} when GANs were introduced in 2014 \citep{2014arXiv1406.2661G}. The first step in the algorithm is to optimize the discriminator $D$. Many researchers initially held the belief that it would be optimal to train $D$ more steps than the generator; however, this is still a controversial topic among researchers \citep{2017arXiv170100160G}. It is Goodfellow's opinion that each model should be trained simultaneously \citep{2017arXiv170100160G}.

>>>>

\section{Introduction}
Deep learning has made astounding progress within the past decade. Discriminative models have already surpassed the abilities of humans to recognize patterns within certain domains \citep{2014arXiv1404.7828S}. These successes can attributed to vast, high dimension datasets in conjunction with large neural networks using linear activation functions, dropout regularization techniques and backpropagation to update their parameters \citep{2014arXiv1406.2661G}. However, deep learning possesses much more ambitious goals within the realm of unsupervised learning. As humans, we are able to understand the world around us with tremendous precision. It is easy to underestimate the complexity of the data we process to accomplish this feat. Although progress within the machine learning field is rapidly advancing, computers maintain a limited understanding of the data they process. Generative models, especially deep generative models such as GANs, offer promising results towards this formidable challenge \citep{genmodelingopenai}.

\subsection{Generative Modeling}
Generative models learn the joint probability distribution $p(x,y)$ of an input $x$ and label $y$ whereas discriminative models directly learn the conditional probability $p(y|x)$. An example of this process is illustrated in figure \ref{fig:pdf}. Generative models may be used as classifiers by applying Bayes rules to calculate the conditional probability $p(y|x)$ \citep{NIPS2001_2020}. However, generative models offer several other applications due to their knowledge of the underlying data distribution. Depending on the model, knowledge of the probability distribution can be created both explicitly and implicitly \citep{Goodfellow-et-al-2016}. Those which do not directly model a probability distribution offer mechanisms which require some implicit knowledge of the underlying distribution, such as generating samples from that distribution \citep{Goodfellow-et-al-2016}.

@@

The process of training a generative model is very similar to a discriminative model. Using large amounts of data collected from a specific domain, we train the generative model to generate data from that domain. As these models typically have fewer parameters than that the number of data samples, they are forced to internalize some representation of the data \citep{genmodelingopenai}. Unlike supervised training, there exists no single desired output. Therefore, the problem becomes defining a cost function which forces the generative model to produce data more like that of the target domain \citep{genmodelingopenai}.

\subsection{Generative Adversarial Networks}
Generative adversarial networks (GANs), first introduced in 2014 by Goodfellow \textit{et al.}, offer a new framework for estimating generative models with use of an adversarial process \citep{2014arXiv1406.2661G}. These models offer a clever approach to solving the aforementioned problem of explicitly defining a cost function. Rather than training a single model, a discriminator is introduced. These two networks are pitted against each other in a minimax, zero-sum game. The generative model $G$ attempts to produce data that resembles that of the training set while the discriminative model $D$ attempts to discriminate between a real and fake (generated) sample. To train this network, $G$ attempts to maximize the error rate of $D$ whereas $D$ attempts to minimize it. The goal of the GAN algorithm is to reduce the distance between the real data and generated data probability distributions. Backpropagation is used to update the parameters of both the descriminator and generator. In the ideal case, $D$ is unable to distinguish generated samples from real samples and produces an error rate of $\frac{1}{2}$ \citep{2014arXiv1406.2661G}.

The adversarial process is often analogized to a counterfeiter trying to fool a detective \citep{2014arXiv1406.2661G}. The detective will receive both real and counterfeited money and will attempt to distinguish between the two. This leads to a competitive scenario where both parties gradually improve their abilities until the counterfeits are indistinguishable from the true currency.

\subsection{Motivations}
There exists several compelling reasons for studying generative modeling, especially GANs \citep{2017arXiv170100160G}. Generative models are extremely useful when your goal is understand the underlying data generating distribution. The ability to learn the joint probability distribution of high dimensional data is relevant to both applied math and engineering, in addition to many other fields \citep{2017arXiv170100160G}. Furthermore, GANs promise numerous real world applications which act as key motivators for their research. Several of these applications are expanded upon in section \ref{sec:applications}.

\subsection{Related Work}
Until the advent of GANs, most deep generative models provided a parametric specification of an probability distribution function \citep{2014arXiv1406.2661G}. These models could be trained using maximum likelihood estimation which often required gradient estimation. The most successful of these models was the deep Boltzmann machine \citep{2014arXiv1406.2661G}. The difficulties associated with gradient estimation motivated the development of \textit{generative machines}. These new models implicitly represent the joint probability distribution by generating samples from that distribution rather than explicitly rendering a probability distribution.

@@

\section{Background}
The goal of the GAN training process is to determine the discriminator and generator parameters such that the classifier error is minimized and maximized respectively \citep{2017arXiv171007035C}. GANs may use any differentiable functions for their underlying implementation; however, optimal results are typically achieved when neural networks are used \citep{2014arXiv1406.2661G}. We first define a vector $z$ which is sampled from a prior noise distribution $z~p_z(z)$. This vector becomes the argument of the generator $G(z;\theta_g)$ where $G$ is a differentiable function and $\theta_g$ are its parameters. Another differentiable function $D(x;\theta_d)$ is defined for the discriminator. This function takes as input a vector $x$ from either the output of $G$ or the training set and is parametrized by $\theta_d$. The output of the discriminator is a scalar number which represents the probability that the sample $x$ was derived from the training data. In this situation, there exists a minimax game, a scenario derived from game theory, where $D$ attempts to maximize the probability of assigning a correct label and $G$ attempts to minimize that same probability. The value function $V(D,G)$ becomes:

@@

This is a standard cross entropy cost function of a neural network that is trained on two minibatches, one from the training dataset and the other from the generator. For this to become a zero-sum game, the cost function of the generator is simply the negation of the discriminator:

@@

This version of the generator loss function is particularly useful for theoretical analysis as shown in the original GAN paper \citep{2014arXiv1406.2661G}. Goodfellow demonstrated that learning within this minimax game is identical to minimizing the Jensen-Shannon divergence and that the game will eventually converge if both players are able to directly update the function space as oppose to the parameters within the network \citep{2017arXiv170100160G}. However, in this minimax version, the loss function of the generator suffers from a vanishing gradient when the discriminator confidently rejects the generated data. The solution proposed by Goodfellow is to train the generator to maximize $log(D(G(z)))$ rather than minimize $log(1-D(G(z)))$ \citep{2014arXiv1406.2661G}. This function is heuristically motivated as to provide strong gradients when either the generator or discriminator is not functioning optimally. In this version, the generator attempts to maximize the log probability of a misclassification whereas the minimax version attempts to minimize the log probability of the discriminator being correct:

@@

A maximum likelihood cost is another option for training GANs \citep{2017arXiv170100160G}. It can be shown that that this cost function minimizes the Kullback-Leibler divergence between $p_{data}$ and $p_{model}$ instead of the Jensen-Shannon divergence. This cost estimator is useful as it allows for comparison with other generative models; however, it behaves similarly to the minimax cost function as seen in figure \ref{fig:CostFunc}. The cost function for this version becomes:

@@

\subsection{Algorithm}
The following algorithm presents the initial training algorithm as introduced by Goodfellow \textit{et al.} in 2014 \citep{2014arXiv1406.2661G}. The first step in the algorithm is to optimize the discriminator $D$. Many researchers initially held the belief that it would be optimal to train $D$ for several iterations before training the generator; however, this is still a controversial topic among researchers \citep{2017arXiv170100160G}. It is Goodfellow's opinion that each model should be trained simultaneously \citep{2017arXiv170100160G}.
